Formal Verification of Real Models from Platforms Using The B-method

### 2009-04-20 papers/FM2009/source/paper.tex

\subsection{Modelling bytes}

@@

$
\begin{array}{l}
\mathit{BYTE\_WIDTH} = 8 \land \mathit{BYTE\_INDEX} = 0..(\mathit{BYTE\_WIDTH}-1) \land \\
\mathit{BYTE} \subseteq \mathit{BIT\_VECTOR} \land \mathit{BYTE} = \{ v \in \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v) = \mathit{BYTE\_WIDTH}\} \land \\
\mathit{BYTE\_ZERO} \in \mathit{BYTE} \land \mathit{BYTE} = \mathit{BYTE\_INDEX} \times \{0\} 
\end{array}
$

@@

\subsection{Manipulation data functions from Z80} [I need review the english] There are some specific functions from
Z80 to manipulate the data. The $\mathit{bv\_ireg\_plus\_d}$ is  used to indexed address. It received the
value of register ($\mathit{ix}$ or $\mathit{iy}$) and displacement to return the sum, the result is the memory address dislocated, see
its the definition.

@@

\begin{itemize}
  \item ALU \it inc  $\in$ \it BYTE $\fun$  \it BYTE \rm - It receives a byte and returns its decrement. There is a
  similar function named $\mathit{dec}$.

  \item ALU \it instruction\_next  $\in$  USHORT  $\fun$  USHORT \rm - It receives the pc actual
  value and return its increment.

  \item ALU \it is\_negative  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns 1 if the received
  byte is zero, otherwise returns 0.

  \item byte \it is\_zero $\in$ BYTE  $\fun$   BIT \rm - It returns 1 if the received byte is zero, otherwise
  returns 0.

  \item byte \it rotateright  $\in$  \it BYTE  $\fun$  \it BYTE \rm - It receives a byte and returns its rotated to
  right. There is a similar function named $\mathit{rotateleft}$.

  \item ALU \it update\_refresh\_reg \rm - It receives a byte and returns its increment until the seventh bit.
\end{itemize}
Every logic functions defined in the $\mathit{BYTE}$ module are included in the $\mathit{ALU}$ module. Then, it was
not necessary rewrite these functions.

@@

\subsubsection{I/O Ports} % Can be placed in basic definitions in a paragraph
%Referencia o modeo de modelagem do captura de um evento externo de acordo com
% o proposto em [DAVID_gdc_2007]

[[ Flag registers and the associated functions: negative, half, zero.. ]]

@@

\hspace*{0.0in}\bf IN\_r\_(C) \rm ( \it rr \rm ) \rm =

>>>>

\subsection{Modelling bytes and bit vector of length 16}[I need review the english]

@@

% $
% \begin{array}{l}
% \mathit{BYTE\_WIDTH} = 8 \land \mathit{BYTE\_INDEX} = 0..(\mathit{BYTE\_WIDTH}-1) \land \\
% \mathit{BYTE} \subseteq \mathit{BIT\_VECTOR} \land \mathit{BYTE} = \{ v \in \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v) = \mathit{BYTE\_WIDTH}\} \land \\
% \mathit{BYTE\_ZERO} \in \mathit{BYTE} \land \mathit{BYTE} = \mathit{BYTE\_INDEX} \times \{0\}
% \end{array}
% $
% Version 2

\hspace*{0.0in}\it BYTE\_WIDTH \rm = 8 $\land$ \it BYTE\_INDEX \rm = 1 $\upto$ \rm (\it BYTE\_WIDTH\rm )$\land$

\hspace*{0.0in}\it PHYS\_BYTE\_INDEX \rm = \rm 0 $\upto$ \rm (\it BYTE\_WIDTH\rm -\rm 1\rm )\hspace*{0.10in} $\land$

\hspace*{0.0in}\it BYTE \rm = \rm \{ \it bt  $\mid$  \it bt $\in$ \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it bt\rm )\rm =\it BYTE\_WIDTH\rm \}\hspace*{0.10in} $\land$

\hspace*{0.0in}\it BYTE\_ZERO  $\in$  \it BYTE  $\land$

\hspace*{0.0in}\it BYTE\_ZERO \rm = \it BYTE\_INDEX  $\times$  \rm \{\rm 0\rm \}

The $\mathit{BYTE\_INDEX}$ is domain of byte modeled. it start at 1 to  obey a definition of sequences from B
method. However, the functions from byte encapsulate the acces and these functions use the
$\mathit{PHYS\_BYTE\_INDEX}$. The $\mathit{BYTE}$ type is a specializated type from $\mathit{BIT\_VECTOR}$, but it
has a size limit. Some other more specific definitions are created to help the developer. For example the constant
$\mathit{BYTE\_ZERO}$ and some below functions.

\begin{itemize}
  \item \it is\_zero $\in$ BYTE  $\fun$   BIT \rm - It returns 1 if the received byte is zero, otherwise
  returns 0.

  \item \it rotateright  $\in$  \it BYTE  $\fun$  \it BYTE \rm - It receives a byte and returns its rotated to
  right. There is a similar function named $\mathit{rotateleft}$.

\end{itemize}

By the very similar way, the type $\mathit{BV16}$ is created.

@@

\begin{property}[Assuring the absence of Overflow]
 To assure that don´t happen overflow the developer can add this expression ($\mathit{bv\_get(rgs8(z0),0)} \neq 1
 \land \mathit{bv\_get(rgs8(z0),2) \neq 1}$)in the invariant. By default, the overflow can happen, but many times it
 can be dangerous. Then, the developer may prohibit its use. This restriction can also become the more dificult to
 specify.
\end{property}

\subsection{Manipulation data functions from Z80} [I need to review the english] There are some specific functions
from Z80 to manipulate the data. The $\mathit{bv\_ireg\_plus\_d}$ is  used to indexed address. It received the
value of register ($\mathit{ix}$ or $\mathit{iy}$) and displacement to return the sum, the result is the memory address dislocated, see
its the definition.

@@

\begin{itemize}
  \item \it inc  $\in$ \it BYTE $\fun$  \it BYTE \rm - It receives a byte and returns its decrement. There is a
  similar function named $\mathit{dec}$.

  \item \it instruction\_next  $\in$  USHORT  $\fun$  USHORT \rm - It receives the pc actual
  value and return its increment.

  \item \it is\_negative  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns 1 if the received
  byte is zero, otherwise returns 0.

  \item \it update\_refresh\_reg \rm - It receives a byte and returns its increment until the seventh bit.

\end{itemize}

Every logic functions defined in the $\mathit{BYTE}$ and $\mathit{BV16}$ module are included in the $\mathit{ALU}$
module. Then, it was not necessary rewrite these functions.

@@

@@

\hspace*{0.0in}\bf IN\_r\_9C0 \rm ( \it rr \rm ) \rm =
